---
title: "MT is Better than AT for Fuel Efficiency (MPG)"
<<<<<<< HEAD
author: "Wenzhe LAI, B.S. of Statistics and Applied Economics"
output: html_document
=======
output: pdf_document
>>>>>>> b9b1d6df36bdddabe5adc338e2231626e67e0a48
---
## Exclusive Summary and Synopsis ##    
This report tries to answer these two questions.     
- *"Is an automatic or manual transmission better for MPG"*      
- *"Quantify the MPG difference between automatic and manual transmissions"*     
By exploring into some data, We found out that the fuel efficiency (Miles/(US) gallon, MPG) is **influnced** by the automaticity of the transmissions system.  
This essay will show all steps during my analysis. All the details will be shown in the 2-page report. In order to make this report reproducible, the codes, graphs and results will be shown on the appendix.

## Part1. Getting and Cleaning Data ##    
In this step, I'll get the dataset `mtcars`. `mtcars` dataset is an embedded dataset in R `datasets` package. It's extracted from the 1974 *Motor Trend* US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles.     
To make the analysis more flexible, we will firstly factorize the discrete variables with `factor` function.  
```{r echo=FALSE}
data(mtcars)
mtcars.fact = mtcars
mtcars.fact$cyl <- factor(mtcars$cyl); mtcars.fact$vs <- factor(mtcars$vs);
mtcars.fact$gear <- factor(mtcars$gear); mtcars.fact$am <- factor(mtcars$am); mtcars.fact$carb <- factor(mtcars$carb);
```

## Part2. Exploratory Data Analysis ##    
<<<<<<< HEAD
In this step, We may take a glimpse of the mtcars data. First use the `cor` function to get the  correlations between `mpg` and each of other variables. Also we will draw a plot Figure 1 of correlations between different variables.  
=======
In this step, We may take a glimpse of the mtcars data. First use the `pairs` function to get the  correlations between `mpg` and each of other variables.   
>>>>>>> b9b1d6df36bdddabe5adc338e2231626e67e0a48
**Table 1. The correlations between the `mpg and other variables**  
```{r echo=FALSE}
round(cor(mtcars$mpg, y = mtcars[, -1]), 3)
```
<<<<<<< HEAD
Second we will draw the box plot Figure 2 of the `mpg` variable against the influence by factor `am`.  
*Figure 1 and Figure 2 will be shown in the Appendix.*      
=======
Second we will draw the box plot of the `mpg` variable against the influence by factor `am`.  
*Figure 1 will be shown in the Appendix.*      
>>>>>>> b9b1d6df36bdddabe5adc338e2231626e67e0a48
As is shown in the *Table 1* and *Figure 1*, we can draw an intuitive conclusion that `am` influences the `mpg` variable. Then we will show and quantify this conclusion.   

## Part3. Inference with the Models ##    
In this part we will analysis deeply into the dataset. Firstly test the `am`'s influence toward the grouped `mpg` means. Secondly find and select the optimal linear regression model.  

### One Way ANOVA, Test of significance of the causality between the `am` and `mpg` ###
ANOVA is used to analyze a factor's influence towards the grouped outputs.
ANOVA is based on the assumption of homogeneity of variances. Let's test it first.
```{r echo=TRUE}
## The P-value is:
bartlett.test(mpg ~ am, data = mtcars.fact)$p.value
```
So we cannot reject the assumption of homogeneity of variances. So we will test the factor `am` with ANOVA next.  
**Table 2. ANOVA Table**
```{r echo=FALSE}
summary(aov(mpg ~ am, data = mtcars.fact))
```
The p-value is significantly small, thus we will draw to the conclusion that the variable `am` influences the mean of different cars' MPG.

### Linear Regression Model Selection ###
In this sub-part, we will firstly fit the linear models for `mpg` against all other variables, and use the `step` function to delect some variable to find the optimal linear models. 
```{r results='hide'}
fit.whole <- lm(mpg ~ ., data = mtcars.fact)
fit.optimal <- step(fit.whole, direction = 'both')
```
```{r}
print(fit.optimal$call)
```
As is printed above, the optimal linear models includes the numeric argument `hp`, `wt`, and factoral argument `cyl`, `am`.  
Now we will test different models with some combinations of arguments `hp` `wt` `cyl` and `am`. We will use the $R^2$ criterion.  
```{r results='hide'}
fit.hpwt <- lm(mpg ~ hp + wt, data = mtcars.fact)
fit.hpwt.cyl <- lm(mpg ~ hp + wt + cyl, data = mtcars.fact)
```
```{r echo=FALSE}
adj.r2 <- c(summary(fit.hpwt)$adj.r.squared, summary(fit.hpwt.cyl)$adj.r.squared, summary(fit.optimal)$adj.r.squared)
names(adj.r2) <- c("hp + wt", "hp + wt + cyl", "hp + wt + cyl + am")
```
**Table 3. The $R^2$ of Each Linear Models**
```{r echo=FALSE}
adj.r2
```
According to the table, the linear model fitted with the variable `am` can fit better, compared to several other models. Thus `am` has the influential effects towards the `mpg`.

## Part4. Diagnostics of the Optimal Linear Models ##    
At the beginning, we will draw some graphs of the optimal regression model. These graphs contains the **Residual vs Fitted Graph**, **Q-Q Graph**, **Scale-Location Graph** and the **Residuals vs Leverage Graph**.    
<<<<<<< HEAD
*Figure 3 is shown on the Appendix*    
=======
*Figure 2 is shown on the Appendix*    
>>>>>>> b9b1d6df36bdddabe5adc338e2231626e67e0a48
Take a glimpse at the Figure 2 four graphs, we can find out that some models is not quite obey the regression model. Obviously, they are **Toyota Corolla** and **Fiat 128**. The Residuals graph shows that the residuals of the two models is quite large, and the Normal Q-Q Plot shows that their residual is doesn't follow the Normal Distribution.   
Regardless of the two special cases, the conclusion that the influence of the `am` towards `mpg` is significant is easy to find out.  


## Appendix ##
  
<<<<<<< HEAD
```{r echo=FALSE}
library(corrplot)
```

###Figure 1 from the Part 2###
```{r}
M <- cor(mtcars)
corrplot.mixed(M, lower = "number", upper = "circle", title = "Correlations between Different Variances")
```
   

###Figure 2 from the Part 2###
**Figure 2. The box plots of the `mpg` variable against the influence by factor `am`**
```{r}
boxplot(mpg ~ am, data = mtcars, names = c("Automatic", "Manual"))
```


###Figure 3 from the Part 4###
**Figure 3. Diagnostical Plots of the Optimal Linear Model**
=======
###Figure 1 from the Part 2###
```{r}
boxplot(mpg ~ am, data = mtcars, names = c("Automatic", "Manual"))
```
**Figure 1. The box plots of the `mpg` variable against the influence by factor `am`**

###Figure 2 from the Part 4###
>>>>>>> b9b1d6df36bdddabe5adc338e2231626e67e0a48
```{r fig.height=7}
par(mfrow = c(2,2))
plot(fit.optimal)
```
<<<<<<< HEAD
    

=======
**Figure 4. Diagnostical Plots of the Optimal Linear Model**

**Table. The Original Dataset of `mtcars`**
```{r echo=FALSE}
mtcars
```
>>>>>>> b9b1d6df36bdddabe5adc338e2231626e67e0a48
